---
title: "Project 2"
author: "Rachael Stryker, Tom Muhlbauer, and Daniel Lassiter"
date: "12/3/2020"
output:
  pdf_document:
    toc: true
    number_sections: true
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    number_sections: true
    theme: united
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, fig.fullwidth=TRUE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(forecast)
library(imputeTS)
library(ggplot2)
library(ggpubr)
library(ggbiplot)
# library(ggfortify)
library(ggResidpanel)
library(MASS)
library(forecast)
library(tidyverse)
library(lubridate)
library(mtsdi)
library("car")
library(tseries)
library(MTS)
library(here)
library(tsibble)
library(feasts)
library(fable)
library(dplyr)
library(timeDate)
sourcedir <- "C:/Users/Daniel/Dropbox/Self organization/Education/UVA/Graduate School/Coursework/Fall_2020/SYS 6021 Linear Statistical Models/Source"
datadir <- "C:/Users/Daniel/Dropbox/Self organization/Education/UVA/Graduate School/Coursework/Fall_2020/SYS 6021 Linear Statistical Models/6021_Project2/"

setwd(sourcedir)

source("SPM_Panel.R")
source("PCAplots.R")
source("FactorPlots.R")
source("pc.glm.R")
source("ROC.R")
source("TestSet.R")


```


```{r, echo = FALSE}
# Load data and impute missing values
setwd(datadir)

airquality = read.csv('AirQualityUCI.csv')

# replace -200 with NA
airquality[airquality == -200] <- NA

# convert integer type to numeric
intcols = c(4,5,7,8,9,10,11,12)
for(i in 1:length(intcols)){
  airquality[,intcols[i]] <- as.numeric(airquality[,intcols[i]])
}

setwd(sourcedir)

# create new data frame with just NO2 and impute missing values
AQdata = airquality["NO2.GT."]
AQdata = na_interpolation(AQdata)

# aggregate to daily maxima for model building
dailyAQ <- aggregate(AQdata, by=list(as.Date(airquality[,1],"%m/%d/%Y")), 
                     FUN=max) %>% as_tsibble(index = Group.1) %>%
          dplyr::rename("Date" = "Group.1") %>% dplyr::rename("NO2" = "NO2.GT.")

# create time series of NO2
NO2.ts <- ts(dailyAQ[,2])
```

# Exploratory Data Analysis
## Inspecting for trends and seasonality
```{r}
# For quickly formatting plots
my_theme <- theme(plot.title = element_text(hjust = 0.5, size = 16), 
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12),
        plot.caption = element_text(size = 12, hjust = 0))
```

```{r}
dailyAQ %>% autoplot(NO2) + ylab("NO2 (mg/m^3)") + xlab("Date") + 
  ggtitle("Nitrogen Concentration") + my_theme
```

The time series plot indicates potential seasonal components and a potential trend component.

```{r}
dailyAQ %>% gg_season(NO2, period = "week") + ylab("NO2 (mg/m^3)") + xlab("") + 
    ggtitle("Nitrogen Concentration") + my_theme
```

This plot indicates an increasing trend over time.


```{r}
dailyAQ %>% model(STL(NO2 ~ trend(window=21) + season(window = "periodic"), robust = TRUE)) %>%
  components() %>%autoplot() + my_theme + ylab("NO2 (mg/m^3)")
```

This seasonal decomposition chart shows a possibly increasing trend and a weekly seasonal component.



```{r}
pg.NO2 <- spec.pgram(NO2.ts,spans=9,demean=T,log='no')

spec.NO2 <- data.frame(freq=pg.NO2$freq, spec=pg.NO2$spec)

spec.NO2 <- data.frame(freq=pg.NO2$freq, spec=pg.NO2$spec) %>% mutate(period = 1/freq) %>% arrange(desc(spec))

spec.NO2[18, ]
```
There is a peak at around 1.5 which corresponds to a seasonality component
with a period of 1 week.
# Need help form Rachael here
  
## Hypotheses from exploratory data analysis:

There is an increasing trend in NO2 emissions. This could be due to increasing cars on the road over time. There is a weekly seasonal component. This could be a function of the number of cars on the road, and perhaps traffic is higher on the week days or weekends.


# Building Univariate Time Series Models
```{r}
# Creating potential predictors
dailyAQ <- dailyAQ %>% mutate(step = 1:length(Date)) %>%
      mutate(day = as.factor(as.character(lubridate::wday(Date, label  = TRUE)))) %>%
      mutate(weekend = isWeekend(Date))

# Create train and test sets
dailyAQ_train <- dailyAQ %>% head(n = length(Date) - 8)
dailyAQ_test <- dailyAQ %>% tail(n = 7)
```


## Modeling trend and season
Deciding how to model the seasonal component was an iterative process. The models shown below include only those we considered to be reasonable candidates. Note that we decided to use the packages feasts, fable, and fabletools for these analyses so our code looks different than the the in class examples.

```{r}
# Create potential trend/season models
models <- dailyAQ_train %>%
  model(trend = TSLM(NO2 ~ trend()),
        trnd_ssn_w_wkday_dummy = TSLM(NO2 ~ trend() + day),         
        trnd_ssn_w_wknd_dummy = TSLM(NO2 ~ trend() + weekend),
        trnd_ssn_wkly = TSLM(NO2 ~ trend() + sin(2*pi*step/7) + cos(2*pi*step/7)), 
        trnd_ssn_wkly_mnthly = TSLM(NO2 ~ trend() + sin(2*pi*step/(365.25/12)) + cos(2*pi*step/(365.25/12))
                                   + sin(2*pi*step/7) + cos(2*pi*step/7)),
        trnd_ssn_complex_trig = TSLM(NO2 ~ trend() + sin(2*pi*step/192) +
                                    cos(2*pi*step/192) + sin(2*pi*step/128) +
                                    cos(2*pi*step/128) + sin(2*pi*step/76.8) +
                                    cos(2*pi*step/76.8)+ sin(2*pi*step/64) +
                                    cos(2*pi*step/64)+ sin(2*pi*step/7) +
                                    cos(2*pi*step/7)+sin(2*pi*step/30) +
                                    cos(2*pi*step/30)+sin(2*pi*step/365) +
                                    cos(2*pi*step/365)),
        trnd_ssn_complex_trig_and_dummy = TSLM(NO2 ~ trend() + weekend + 
                                    sin(2*pi*step/192) + 
                                    cos(2*pi*step/192) + sin(2*pi*step/128) + 
                                    cos(2*pi*step/128) + sin(2*pi*step/76.8) +
                                    cos(2*pi*step/76.8)+ sin(2*pi*step/64) +
                                    cos(2*pi*step/64)+ sin(2*pi*step/7) +
                                    cos(2*pi*step/7)+sin(2*pi*step/30) +
                                    cos(2*pi*step/30)+sin(2*pi*step/365) +
                                    cos(2*pi*step/365))
        )

```


```{r}
# Inspect the trend only model
models %>% dplyr::select(trend) %>% report()
```
The coefficient on trend is significant, and plotting the modeled trend line 
against the original data below seems to verify this. The significance of the trend will change when the seasonal components are modeled but it turns out to be significant in all cases.

```{r}
models %>% dplyr::select(trend) %>% fitted() %>%
  autoplot(color = 'red') + autolayer(dailyAQ_train) +
        ggtitle("Trend Model") + ylab("NO2 mg/m^3") + xlab("") + my_theme
```

Next, we inspected the reports and plots of the trend season models that used day-of-the-weekend and weekend/not weekend dummy variables.


```{r}
models %>% dplyr::select(trnd_ssn_w_wkday_dummy) %>% report()

models %>% dplyr::select(trnd_ssn_w_wkday_dummy) %>% fitted() %>%
  autoplot(color = 'red') + autolayer(dailyAQ_train) +
        ggtitle("Trend-Season Model with Day-of-the-week Dummy Variables") + ylab("NO2 mg/m^3") + xlab("") + my_theme
```


This model used a dummy variable for each day of the week. Friday is the base case.
Only Saturday and Sunday were significantly different from the base case. NO2
concentrations on these days are lower. Perhaps traffic is heavier during the
weekdays because of commuters. It is also notable that the trend term is still
significant once the seasonality component has been accounted for.

The insignificance of the weekday variables led us to attempt to model the seasonal component with a single dummy variable 
indicating whether the date was on a weekday or weekend.

```{r}
models %>% dplyr::select(trnd_ssn_w_wknd_dummy) %>% report()

models %>% dplyr::select(trnd_ssn_w_wknd_dummy) %>% fitted() %>%
  autoplot(color = 'red') + autolayer(dailyAQ_train) +
        ggtitle("Trend-Season Model with Dummy Variables") + ylab("NO2 mg/m^3") + xlab("") + my_theme
```

All coefficients are significant indicating this is a strong candidate model.


Next we inspected the model reports and plot of the trend season models that relied on trigonometric functions. We show this all in a single graph and table to save space. Note that the model 'trnd_ssn_complex_trig_and_dummy' does in fact have a dummy variable indicating whether a day is a weekend or not. The table showing coefficients and p-values are sorted from highest p-value to lowest p-value.


```{r, fig.height=6, fig.width=8, fig.align='center'}
models %>% dplyr::select(trnd_ssn_wkly, trnd_ssn_wkly_mnthly, 
                         trnd_ssn_complex_trig,
                         trnd_ssn_complex_trig_and_dummy) %>% coef() %>% arrange(desc(p.value))

models %>% 
  dplyr::select(trnd_ssn_wkly, trnd_ssn_wkly_mnthly, 
                trnd_ssn_complex_trig, trnd_ssn_complex_trig_and_dummy) %>% 
  fitted() %>% autoplot() +
  autolayer(dailyAQ_train) + 
  ggtitle("Trend-Season Model with Trigonometric Functions") + 
  ylab("NO2 mg/m^3") + xlab("") + my_theme
```

Visually, the complex trig functions track the process very well though there are a few insignificant variables. Interestingly, the cosine function counterpart to the most insignificant variable of the trnd_ssn_complex_trig model (the sine function component of a 76.8 day season) has a p-value significant to the 0.001 level. This is not shown in the table but could easily be recreated from the attached .rmd.

To decide which is the best approach, we compared performance metrics across
the candidate models with a trend and seasonal component.

```{r}

models %>% 
  dplyr::select(trnd_ssn_wkly, trnd_ssn_wkly_mnthly, trnd_ssn_complex_trig,
                trnd_ssn_complex_trig_and_dummy,
         trnd_ssn_w_wkday_dummy, trnd_ssn_w_wknd_dummy) %>% 
  report() %>% dplyr::select(.model, adj_r_squared, AIC, BIC, df.residual)
```
The trend/season models with the complex trigonometric functions clearly outperformed the other models, but we decided to include other model candidates in case these are overfit. Since the only difference between the trnd_ssn_wkly and trnd_ssn_wkly_mnthly model is the inclusion of a monthly trigonometric function and the latter performed better, we decided to exclude the trnd_ssn_wkly model from the rest of our analysis.


## Modeling the residuals of best trend/season models
Next, we extracted and plotted the residuals of each candidate trend/season model to inspect for autoregressive properties.

```{r, fig.height=6, fig.width=8, fig.align='center'}
# Extract model residuals
data_resids <- models %>%
  dplyr::select(trnd_ssn_wkly_mnthly, trnd_ssn_w_wkday_dummy,
                trnd_ssn_w_wknd_dummy, trnd_ssn_complex_trig,
                trnd_ssn_complex_trig_and_dummy) %>% 
  residuals()

# Plot the residuals
data_resids %>% autoplot(.resid) + ggtitle("Trend/Season Model Residuals") + 
  xlab("") +  my_theme
```

There is clearly some autocorrelation in the residuals in all of the models because none of them look like random white noise. The next step was to inspect the ACF and PACF plots to decide how to model them. Because there are so many models, we look at the dummy variable models separate from the trigonometric models.

```{r, fig.height=11, fig.width=8, fig.align='center'}
plt_resid_ACF <- data_resids %>%
  dplyr::filter(.model %in% c("trnd_ssn_w_wknd_dummy" ,"trnd_ssn_w_wkday_dummy")) %>%
  ACF(.resid) %>% autoplot() + ggtitle("Dummies: NO2 Residuals ACF plot") + my_theme

plt_resid_PACF <- data_resids %>%
  dplyr::filter(.model %in% c("trnd_ssn_w_wknd_dummy" ,"trnd_ssn_w_wkday_dummy")) %>%
  PACF(.resid) %>% autoplot() + ggtitle("Dummies: NO2 Residuals PACF plot") + my_theme

ggarrange(plt_resid_ACF,plt_resid_PACF,nrow=2,ncol=1)
```

Both dummy variable trend/season models exhibit similar behavior in the ACF and PACF. 
There is sinusoidal decay in the ACF of the residuals. The PACF could be
interpreted as  cutting off after lag 2 or as exhibiting sinusoidal behavior. 
This indicates that the residuals could be modeled as either AR(2) or as an
ARMA process.


```{r, fig.height=11, fig.width=8, fig.align='center'}
 plt_resid_ACF <- data_resids %>%
  dplyr::filter(.model %in% c("trnd_ssn_wkly_mnthly" ,"trnd_ssn_complex_trig", "trnd_ssn_complex_trig_and_dummy")) %>%
  ACF(.resid) %>% autoplot() + ggtitle("Trig Functions: NO2 Residuals ACF plot") + my_theme

plt_resid_PACF <- data_resids %>%
  dplyr::filter(.model %in% c("trnd_ssn_wkly_mnthly" ,"trnd_ssn_complex_trig", "trnd_ssn_complex_trig_and_dummy")) %>%
  PACF(.resid) %>% autoplot() + ggtitle("Trig Functions: NO2 Residuals PACF plot") + my_theme

ggarrange(plt_resid_ACF,plt_resid_PACF,nrow=2,ncol=1)
```

The two trend/season models with more trigonometric functions (denoted 'complex') potentially show sinusoidal behavior in the ACF and the PACF and the PACF appears to cut off after 1 lag, although the second lag is technically barely above the significant threshold for the one that includes a weekend dummy variable. This indicates that it these should either be modeled as an AR(1) process or possibly an ARMA process if the PACF could be considered to be sinusoidal. The less complex trend/season model that only has monthly and weekly components looks more like the dummy variable models above with sinusoidal decay in the ACF and a PACF that either cuts off at lag 2, indicating AR(2), or a sinuisoidal PACF which would lead to modeling it as an ARMA process.

Next we performed autoregressive modeling of the residuals according to these hypotheses. Because of the structure of the tibble package, we had to apply the AR(1) and AR(2) models to all residual datasets. We also deployed an auto-selection process.

```{r}
# Model the residuals
models_residuals <- data_resids %>%
  rename("model" = ".model", "resid" = ".resid") %>%
  model(auto.arima = ARIMA(resid ~ PDQ(0,0,0)), 
        ar2 = ARIMA(resid ~ pdq(2, 0, 0) + PDQ(0,0,0)),
        ar1 = ARIMA(resid ~ pdq(1, 0, 0) + PDQ(0,0,0)))

# Show the pdq values from the auto.arima models
models_residuals %>% tidy() %>% filter(.model == 'auto.arima') %>% print(n = 3e3)
```

The auto-selection process yielded ARMA(2, 1) models for the day-of-the-week and weekend/not weekend dummy variable models as well as the weekly and monthly trigonometric models. The auto-selection process yielded an AR(1) model for the complex_trig models as we hypothesized based on the ACF and PACF.

Next we report the AIC and BIC for each of the candidate residual models.

```{r}
# Compare AIC and BIC across the models; sort by AIC
models_residuals %>% report() %>%
  dplyr::select(model, .model, AIC, BIC) %>% arrange(AIC) %>% print(n = 100)

```

Based on AIC, the auto-selected ARIMA model performed best or tied for the best model for each set of residuals. BIC did not align exactly with AIC results. For both the dummy variable trend/season model residuals, the AR(2) model had the best score which was our hypothesis based on the AIC and BIC. The score improvement is likely negligable. For ease of visualization, we decided to continue with the rest of our analysis using only the auto-selected models.

Next we plotted residuals versus fitted and QQ plots of model residuals to test for the iid assumption.

```{r, fig.height=11, fig.width=8, fig.align='center'}
# Choose the auto.arima model for the rest of the analysis
models_residuals <- models_residuals %>% dplyr::select(-ar2, -ar1)

# Plot residuals vs. fitted
models_residuals_fitted <-
  models_residuals %>% fitted() %>% as_tibble() %>% dplyr::select(Date, '.fitted')

models_residuals_resid <- 
  models_residuals %>% residuals() %>% as_tibble()

model_residuals_fittedAndResiduals <- 
  inner_join(models_residuals_resid, models_residuals_fitted, by = "Date")

ggplot(model_residuals_fittedAndResiduals, aes(.fitted, .resid)) +
    geom_point() + facet_grid(rows = vars(model), cols = vars(.model))
```

There is no clear trend in any of the model residuals.

```{r, fig.height=11, fig.width=8, fig.align='center'}
# Plot QQ plots
ggplot(model_residuals_fittedAndResiduals, aes(sample = .resid)) +
    stat_qq() + stat_qq_line(color = "red") + 
    facet_grid(rows = vars(model), cols = vars(.model))
```

The residuals for each model all appear to exhibit similar patterns and are 
approximately gaussian with some fat-tail behavior.

The next step was to forecast and test each model against the test set.

# Forecasting

We started by forecasting the residuals.

```{r, fig.height=11, fig.width=8, fig.align='center'}
resid_summary <- models_residuals %>% augment()

resid_7day_forecast <- models_residuals %>%
  forecast(h = "1 week")

plt_resid_forecast <- resid_7day_forecast %>% feasts::autoplot(alpha = 0.8) +
  geom_line(data = resid_summary, aes(x = Date, y = resid)) + 
  ggtitle("Forecast of residuals") + my_theme

plt_resid_forecast

resid_7day_forecast <- resid_7day_forecast %>% hilo(level = 95)

```

The forecast of residuals look reasonable and very similar as does the width of the confidence intervals. Next we forecasted the trend/season models.


```{r, , fig.height=11, fig.width=8, fig.align='center'}
main_summary <- models %>% select(trnd_ssn_w_wknd_dummy,
                                  trnd_ssn_w_wkday_dummy,
                                  trnd_ssn_complex_trig_and_dummy,
                                  trnd_ssn_complex_trig,
                                  trnd_ssn_wkly_mnthly) %>% augment()

main_7day_forecast <- models %>% 
  select(trnd_ssn_w_wknd_dummy, trnd_ssn_w_wkday_dummy, 
         trnd_ssn_complex_trig_and_dummy, 
         trnd_ssn_complex_trig, trnd_ssn_wkly_mnthly) %>%
  forecast(new_data = dailyAQ_test)

plt_main_forecast <- main_7day_forecast %>% feasts::autoplot(alpha = 0.8) +
  geom_line(data = main_summary, aes(x = Date, y = NO2)) + 
  ggtitle("Forecast of Trend-Season Model") + 
  facet_grid(rows = vars(.model)) +
  my_theme

plt_main_forecast
```
The trend/season forecasts also look similar with similar confidence intervals. The next step was to combine these models to form the full forecast and to evaluate performance. We show the mean squared error for each model after the plot below.


## Combining Season/trend and residuals forecasts and computing MSE
```{r}
full_model_summary <- resid_summary %>% 
  left_join(main_summary,
            by = c("Date" = "Date", "model" = ".model"),
            suffix = c(".resid", ".full"))

full_forecast <- left_join(resid_7day_forecast, main_7day_forecast,
                           by = c("model" = ".model", "Date" = "Date"),
                           suffix = c(".resid", ".full")) %>%
                mutate(pt_pred = .mean.resid + .mean.full) 

test_data <- dailyAQ_test %>%
  dplyr::select(Date, "NO2") %>%
  rename("NO2_Actual" = "NO2")

full_forecast <- full_forecast %>% inner_join(test_data, by = ("Date" = "Date"))

forecast_MSE <- full_forecast %>% as_tibble() %>% group_by(model, .model) %>%
  mutate(sqrd_error = (pt_pred - NO2_Actual)^2) %>% summarize(MSE = mean(sqrd_error))

```


## Ploting forecasts and MSE
```{r, fig.height=5.5, fig.width=10, fig.align='center'}
full_forecast <- full_forecast %>% unpack_hilo(cols = "95%") %>%
  mutate(`95%_upper` = .mean.full + `95%_upper`) %>%
  mutate(`95%_lower` = .mean.full + `95%_lower`) %>% 
  mutate(model_ID = paste(model, .model, sep = "_"))
  
plot_forecasts <- ggplot(data = full_forecast, aes(group = model_ID, color = model_ID)) + 
  geom_line(aes(x=Date,y=NO2_Actual), color = "black") + 
  geom_line(aes(x=Date,y=pt_pred)) + 
  geom_line(aes(x=Date,y=`95%_lower`),linetype="dashed") + 
  geom_line(aes(x=Date,y=`95%_upper`),linetype="dashed") +
  xlab("") + ylab("NO2 mg/m^3") + 
  ggtitle("NO2 Season/Trend Model + ARMA of Residuals with 95% CIs") + my_theme

plot_forecasts

forecast_MSE
```

Despite being significantly outperformed based on AIC, BIC, adjust R^2, and residuals, the model with the trigonometric functions representing weekly and monthly cycles performed best on the test dataset. That is the model we have selected. This indicates that the models with many trigonometric functions were overfitted.

Next we used the best model to simulate the next year of data.

# Simulation with best model

We started by forecasting the residuals.
```{r}
resid_1year_simulation <- models_residuals %>% dplyr::select(auto.arima) %>%
  filter(model == "trnd_ssn_wkly_mnthly") %>%
  forecast(h = "1 year") # %>% hilo(level = 95)

resid_1year_simulation <- models_residuals %>% dplyr::select(auto.arima) %>%
  filter(model == "trnd_ssn_wkly_mnthly") %>% generate(h = "1 year", seed = 1)
```

Next we forecasted the best trend/season model.
```{r}
vec_simulation_dates <- seq(ymd('2005-03-29'),ymd('2006-03-29'),by='days')

original_steps <- length(dailyAQ_train$Date)
sim_steps <- length(vec_simulation_dates)

simulation_dates <- as_tibble() %>% as_tibble(.rows = length(vec_simulation_dates)) %>%
  bind_cols(vec_simulation_dates) %>% rename('Date' = '...1') %>% 
  mutate(NO2 = -999) %>%
  mutate(step = (original_steps+1):(original_steps + sim_steps)) %>%
  mutate(day = as.factor(as.character(lubridate::wday(Date, label  = TRUE)))) %>%
  mutate(weekend = isWeekend(Date)) %>% as_tsibble(index = Date)


main_1year_forecast <- models %>% 
  select(trnd_ssn_wkly_mnthly) %>%
  forecast(new_data = simulation_dates)

```


Finally, we combined them into a single simulation.
```{r}
full_simulation <- left_join(resid_1year_simulation, main_1year_forecast,
                           by = c("model" = ".model", "Date" = "Date"),
                           suffix = c(".resid", ".full")) %>%
                mutate(pt_pred = .sim + .mean)  %>% 
                mutate(model_ID = paste(model, .model, "_"))
```



```{r, fig.height=5.5, fig.width=10, fig.align='center'}

dailyAQ_train <- dailyAQ_train %>% mutate(model_ID = "Training Data")

plot_simulation <- ggplot(data = full_simulation, aes(group = model_ID, color = model_ID)) + 
  geom_line(aes(x=Date,y=pt_pred)) + 
  geom_line(data = dailyAQ_train, aes(x = Date, y = NO2)) +
  xlab("") + ylab("NO2 mg/m^3") + 
  ggtitle("NO2 Season/Trend Model + ARMA of Residuals with 95% CIs") + my_theme

plot_simulation
```

The final simulation looks reasonable, so we are comfortable reporting this model as our final product.

