---
title: "Project 2"
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    number_sections: true
    theme: united
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, fig.fullwidth=TRUE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(forecast)
library(imputeTS)
library(ggplot2)
library(ggpubr)
library(ggbiplot)
library(ggfortify)
library(ggResidpanel)
library(MASS)
library(forecast)
library(tidyverse)
library(lubridate)
library(mtsdi)
library("car")
library(tseries)
library(MTS)
library(here)
library(tsibble)
library(feasts)
library(fable)

sourcedir <- "C:/Users/Daniel/Dropbox/Self organization/Education/UVA/Graduate School/Coursework/Fall_2020/SYS 6021 Linear Statistical Models/Source"
datadir <- paste0(here(), "/")

setwd(sourcedir)

source("SPM_Panel.R")
source("PCAplots.R")
source("FactorPlots.R")
source("pc.glm.R")
source("ROC.R")
source("TestSet.R")


```


# Load data and impute missing values
```{r}
setwd(datadir)

airquality = read.csv('AirQualityUCI.csv')

# replace -200 with NA
airquality[airquality == -200] <- NA

# convert integer type to numeric
intcols = c(4,5,7,8,9,10,11,12)
for(i in 1:length(intcols)){
  airquality[,intcols[i]] <- as.numeric(airquality[,intcols[i]])
}

setwd(sourcedir)

# create new data frame with just NO2 and impute missing values
AQdata = airquality["NO2.GT."]
AQdata = na_interpolation(AQdata)

# aggregate to daily maxima for model building
dailyAQ <- aggregate(AQdata, by=list(as.Date(airquality[,1],"%m/%d/%Y")), 
                     FUN=max) %>% as_tsibble(index = Group.1) %>%
          rename("Date" = "Group.1") %>% rename("NO2" = "NO2.GT.")

# create time series of NO2
NO2.ts <- ts(dailyAQ[,2])
```

# Exploratory Data Analysis
```{r}
# For quickly formatting plots
my_theme <- theme(plot.title = element_text(hjust = 0.5, size = 16), 
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12),
        plot.caption = element_text(size = 12, hjust = 0))
```

```{r}
dailyAQ %>% autoplot(NO2) + ylab("NO2 (mg/m^3)") + xlab("Date") + 
  ggtitle("Nitrogen Concentration") + my_theme
```

The time series plot indicates potential seasonal components and a potential trend component.

```{r}
dailyAQ %>% gg_season(NO2, period = "week") + ylab("NO2 (mg/m^3)") + xlab("") + 
    ggtitle("Nitrogen Concentration") + my_theme
  
```

This plot indicates an increasing trend over time.


```{r}
dailyAQ %>% model(STL(NO2 ~ trend(window=21) + season(window = "periodic"), robust = TRUE)) %>%
  components() %>%autoplot() + my_theme + ylab("NO2 (mg/m^3)")
```

This seasonal decomposition chart shows a possibly increasing trend and a weekly seasonal component.


```{r}
dailyAQ %>% ACF(NO2) %>% autoplot() + ggtitle("Nitrogen Concentration ACF plot") + my_theme

dailyAQ %>% PACF(NO2) %>% autoplot() + ggtitle("Nitrogen Concentration PACF plot") + my_theme
```

* The ACF shows sinusoidal decay. 
  * This indicates that we need to include an AR term in the model.
* The PACF decreases from the 1st to the 2nd lag and becomes insignificant at the 4th lag. 
  * This indicates that our AR component should be order 3. (?)
* Significant lags also occur after the 4th lag.
* The plot may show some seasonality.
  * If this is the case, then there would be a moving average term. However, 
  the ACF doesn't cut off so it's likely it is an AR(3) model only.


```{r}
pg.NO2 <- spec.pgram(NO2.ts,spans=9,demean=T,log='no')

spec.NO2 <- data.frame(freq=pg.NO2$freq, spec=pg.NO2$spec)
ggplot(spec.NO2) + geom_line(aes(x=freq,y=spec)) + 
  ggtitle("Smooth Periodogram of NO2") + my_theme

spec.NO2 <- data.frame(freq=pg.NO2$freq, spec=pg.NO2$spec) %>% mutate(period = 1/freq) %>% arrange(desc(spec))
spec.NO2[18, ]
```
* There is a peak at around 1.5 which corresponds to a seasonality component
  with a period of 1 week.
  
## Hypotheses from exploratory data analysis:

* There is an increasing trend in NO2 emissions. This could be due to increasing cars on the road over time.
* There is a weekly seasonal component. This could be a function of the number of cars on the road,
  and perhaps traffic is higher on the weekends.
* The ACF and PACF plots indicate that the process could be autorgressive with order 3.



# Building Univariate Time Series Models
```{r}
# Remove last 7 days of data
dailyAQ_train <- dailyAQ %>% head(n = length(Date) - 7)
```


## Modeling
```{r}
library(timeDate)


dailyAQ_train <- dailyAQ_train %>% mutate(step = 1:length(Date)) %>%
      mutate(day = as.factor(as.character(lubridate::wday(Date, label  = TRUE)))) %>%
      mutate(weekend = isWeekend(Date))

models <- dailyAQ_train %>%
  model(trend = TSLM(NO2 ~ trend()),
        trend_season_7 = TSLM(NO2 ~ trend() + sin(2*pi*step/7) + cos(2*pi*step/7)), 
        season_7 = TSLM(NO2 ~ sin(2*pi*step/7) + cos(2*pi*step/7)),
        trend_season_30 = TSLM(NO2 ~ trend() + sin(2*pi*step/30) + cos(2*pi*step/30)), 
        season_30 = TSLM(NO2 ~ sin(2*pi*step/30) + cos(2*pi*step/30)),
        trend_season_w_wkday_dummy = TSLM(NO2 ~ trend() + day),         
        trend_season_w_wknd_dummy = TSLM(NO2 ~ trend() + weekend))
```


### Trend
```{r}
models %>% select(trend) %>% report()
```
The coefficient on trend is significant, and plotting the modeled trend line 
against the original data below verifies this.

```{r}
models %>% select(trend) %>% fitted() %>%
  autoplot(color = 'red') + autolayer(dailyAQ_train) +
        ggtitle("Trend Model") + ylab("NO2 mg/m^3") + xlab("") + my_theme
```


### Seasonality
The model summary and graph of the auto-selected trend and season model 
shows that the trend is still significant. The 3 week and 4 week coefficients


```{r}
models %>% select(trend_season_w_wkday_dummy) %>% report()

models %>% select(trend_season_w_wkday_dummy) %>% fitted() %>%
  autoplot(color = 'red') + autolayer(dailyAQ_train) +
        ggtitle("Trend-Season Model with Dummy Variables") + ylab("NO2 mg/m^3") + xlab("") + my_theme
```


This model used a dummy variable for each weekday. Friday is the base case.
Only Saturday and Sunday were significantly different from the base case. NO2
concentrations on these days are lower. Perhaps traffic is heavier during the
weekdays because of commuters. It is also notable that the trend term is still
significant once the seasonality component has been accounted for.

The models reported in the summary and plot below had trigonometric functions 
to model 7 day and 30 day seasonal components.
```{r}
models %>% select(trend_season_7, trend_season_30) %>% coef()

models %>% 
  select(trend_season_7, trend_season_30) %>% 
  fitted() %>% autoplot() +
  autolayer(dailyAQ_train) + 
  ggtitle("Trend-Season Model with Trigonometric Functions") + 
  ylab("NO2 mg/m^3") + xlab("") + my_theme
```

To decide which is the best approach, we compared performance metrics across
the models with a trend and seasonal component.

```{r}
models %>% 
  select(trend_season_7, trend_season_30, trend_season_w_wkday_dummy) %>% 
  report() %>% select(.model, adj_r_squared, AIC, BIC, df.residual)
```

The third model in the list which was built with day-of-the-week dummy variables
has the best adjust R-squared, AIC, and the lowest residuals. The model built
with trigonometric functions and a 7 day wavelength had a better BIC, so either
could be a valid way to model seasonality.

Next I tried modeling the seasonal component with a single dummy variable 
for whether the date was on a weekday or weekend.

```{r}
models %>% select(trend_season_w_wknd_dummy) %>% report()

models %>% select(trend_season_w_wknd_dummy) %>% fitted() %>%
  autoplot(color = 'red') + autolayer(dailyAQ_train) +
        ggtitle("Trend-Season Model with Dummy Variables") + ylab("NO2 mg/m^3") + xlab("") + my_theme
```

Unsurprisingly, the variable was significant. Finally, to compare with the rest:

```{r}
models %>% 
  select(trend_season_7, trend_season_30, 
         trend_season_w_wkday_dummy, trend_season_w_wknd_dummy) %>% 
  report() %>% select(.model, adj_r_squared, AIC, BIC, df.residual)

```


The model with the weekend/weekdaay dummy variable performed best across all 
metrics except for the residuals, but the difference there is small. This is
how I will model seasonality for the remainder of this analysis.