---
title: "Project 2"
author: "Daniel, Tom, and Rachael"
date: "Date"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
require("knitr")
#sourcedir <- "D:/GoogleDrive/Julie_SYS4021/Projects/TimeSeries2020"
#datadir <- "D:/GoogleDrive/Julie_SYS4021/Data Sets/AirQualityUCI"

datadir <- "C:/Users/tcmuh/Desktop/Fall2020/SYS6021/Data/AirQualityUCI"
sourcedir <-"C:/Users/tcmuh/Desktop/Fall2020/SYS6021/Source"

opts_knit$set(root.dir = sourcedir)
library(forecast)
library(imputeTS)
library(mtsdi)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(ggfortify)
library(ggpubr)
library(tseries)
```

# Load data and impute missing values
```{r cars, warning=FALSE}
setwd(datadir)

airquality = read.csv('AirQualityUCI.csv')

# replace -200 with NA
airquality[airquality == -200] <- NA

# convert integer type to numeric
intcols = c(4,5,7,8,9,10,11,12)
for(i in 1:length(intcols)){
  airquality[,intcols[i]] <- as.numeric(airquality[,intcols[i]])
}

setwd(sourcedir)

# create new data frame with just NO2 and impute missing values
AQdata = airquality["NO2.GT."]
AQdata = na_interpolation(AQdata)

# aggregate to daily maxima for model building
dailyAQ <- aggregate(AQdata, by=list(as.Date(airquality[,1],"%m/%d/%Y")), FUN=max)

# create time series of NO2
NO2.ts <- ts(dailyAQ[,2])
```

First need to visualize the data:

```{r}

AQdata.ts <- ts(AQdata)
autoplot(AQdata.ts)

dailyAQ.train <- dailyAQ[1:(dim(dailyAQ)[1]-7),]

dailyAQ.ts <- ts(dailyAQ.train$NO2.GT.)
autoplot(dailyAQ.ts)

```



```{r}

spec.pgram(dailyAQ.ts,spans=9,demean=T,log='no')


```

From PG highest peak at 192 days, next one may be about 7 days.

Also looks like it may have a constant mean for the first 180-200 days and then an increasing trend.

Next step is to look at the 7-day possible trend.

Also decided to try to look at the 192-day possible trend.

Finally, also decided to put the 7 day and 192 day together on one plot.

```{r}

time.NO2 <- c(1:(length(dailyAQ.ts)))
NO2.ts.lm1 <- lm(dailyAQ.train$NO2.GT.~time.NO2+sin(2*pi*time.NO2/7)+cos(2*pi*time.NO2/7))

NO2.ts.lm2 <- lm(dailyAQ.train$NO2.GT.~time.NO2+sin(2*pi*time.NO2/192)+cos(2*pi*time.NO2/192))

NO2.ts.lm3 <- lm(dailyAQ.train$NO2.GT.~time.NO2+sin(2*pi*time.NO2/7)+cos(2*pi*time.NO2/7)+sin(2*pi*time.NO2/192)+cos(2*pi*time.NO2/192))

summary(NO2.ts.lm1)
summary(NO2.ts.lm2)
summary(NO2.ts.lm3)

ggplot(dailyAQ.train[1:(length(dailyAQ.ts)),], aes(x=time.NO2,y=NO2.GT.)) + geom_line() +   geom_line(aes(x=time.NO2,y=NO2.ts.lm1$fitted.values),color="red") + xlab("") + ylab("NO2")

ggplot(dailyAQ.train[1:(length(dailyAQ.ts)),], aes(x=time.NO2,y=NO2.GT.)) + geom_line() +   geom_line(aes(x=time.NO2,y=NO2.ts.lm2$fitted.values),color="red") + xlab("") + ylab("NO2")

ggplot(dailyAQ.train[1:(length(dailyAQ.ts)),], aes(x=time.NO2,y=NO2.GT.)) + geom_line() +   geom_line(aes(x=time.NO2,y=NO2.ts.lm3$fitted.values),color="red") + xlab("") + ylab("NO2")

```

It looks like both of the seasonal trends may be descriptive of the time series data
since at least one of the coefficients of the sin or cos term is valid at the
0.05 level.

Next need to look at some diagnostics for these three models.

```{r}

AIC(NO2.ts.lm1)
AIC(NO2.ts.lm2)
AIC(NO2.ts.lm3)

```

As an initial look, the AIC for the time series including both the 7-day and 
192-day seasons is the best so we should continue with this model until we
find something that works better.

Maybe add a 10-day season as well?

```{r}

NO2.ts.lm4 <- lm(dailyAQ.train$NO2.GT.~time.NO2+sin(2*pi*time.NO2/10)+cos(2*pi*time.NO2/10))

summary(NO2.ts.lm4)

ggplot(dailyAQ.train[1:(length(dailyAQ.ts)),], aes(x=time.NO2,y=NO2.GT.)) + geom_line() +   geom_line(aes(x=time.NO2,y=NO2.ts.lm4$fitted.values),color="red") + xlab("") + ylab("NO2")

```

10 days was not useful in describing the trend.

Not sure whether this is a normal technique or not but what about looking at a
lot of different values for this?  Prime numbers are probably the best to look at
first.

```{r}

i <- 364

1/i

NO2.ts.lm5 <- lm(dailyAQ.train$NO2.GT.~time.NO2+sin(2*pi*time.NO2/i)+cos(2*pi*time.NO2/i))
  
summary(NO2.ts.lm5)
AIC(NO2.ts.lm5)

ggplot(dailyAQ.train[1:(length(dailyAQ.ts)),], aes(x=time.NO2,y=NO2.GT.)) + geom_line() +   geom_line(aes(x=time.NO2,y=NO2.ts.lm5$fitted.values),color="red") + xlab("") + ylab("NO2%%")

```

28 (AIC 3982.339) and 64 (AIC 3970.735) both look interesting

128: AIC 3988.009
77: AIC 3980.565
64*7=448: AIC 3959.162
364: AIC 3956.527
365: AIC 3956.562

Looks like 365 is a good choice. Next we'll build a model with the three timeframes
7, 192, and 365.

```{r}

NO2.ts.lm6 <- lm(dailyAQ.train$NO2.GT.~time.NO2+sin(2*pi*time.NO2/7)+cos(2*pi*time.NO2/7)+sin(2*pi*time.NO2/192)+cos(2*pi*time.NO2/192)+sin(2*pi*time.NO2/365)+cos(2*pi*time.NO2/365))

summary(NO2.ts.lm6)
AIC(NO2.ts.lm6)

ggplot(dailyAQ.train[1:(length(dailyAQ.ts)),], aes(x=time.NO2,y=NO2.GT.)) + geom_line() +   geom_line(aes(x=time.NO2,y=NO2.ts.lm6$fitted.values),color="red") + xlab("") + ylab("NO2")

```

The AIC dropped significantly for this model. Maybe try it with just 7-day and 
365-day seasons to see if that is just as good.

```{r}

NO2.ts.lm7 <- lm(dailyAQ.train$NO2.GT.~time.NO2+sin(2*pi*time.NO2/7)+cos(2*pi*time.NO2/7)+sin(2*pi*time.NO2/365)+cos(2*pi*time.NO2/365))

summary(NO2.ts.lm7)
AIC(NO2.ts.lm7)

ggplot(dailyAQ.train[1:(length(dailyAQ.ts)),], aes(x=time.NO2,y=NO2.GT.)) + geom_line() +   geom_line(aes(x=time.NO2,y=NO2.ts.lm7$fitted.values),color="red") + xlab("") + ylab("NO2")

```

AIC was better with all three seasonal terms (7-, 192-, and 365-day). We can explain
the 7- and 365-day intervals but the 192-day is a mystery (it is one half the length
of the data set of 384 observations). It is possible that this will lead to 
overfitting the data and may perform worse on the test data.

Recommend we look at both the 7 and 365-day seasons for sure and compare the results
to the model using the 192-day season as well.

Lets take a look at the residuals now.

```{r}

plot(NO2.ts.lm6$residuals)

```

The residuals don't look too bad although they may be growing slightly in magnitude
from the start to the end. I think they appear to be zero-mean.

Next step is to model the residuals.

```{r}

AQresiduals <- NO2.ts.lm6$residuals

AQresiduals.ts <- ts(AQresiduals)

autoplot(AQresiduals.ts)

```

Is there a pattern to try to model with the residuals? I think so but let's look
at the ACF and PACF and see what might be in there.

```{r}

AQres.acf <- ggAcf(AQresiduals.ts)
AQres.pacf <- ggPacf(AQresiduals.ts)
ggarrange(AQres.acf,AQres.pacf,nrow=2,ncol=1)

```

From the ACF, the residuals may not be stationary. From the PACF, there are many
that are significant: at 1, 3, 7, 12, and 15 (not sure about 2).

The next step is to do an ARIMA model and see what comes out.

```{r}

AQres.auto <- auto.arima(AQresiduals.ts,approximation=FALSE)
summary(AQres.auto)

```

So no difference was needed, only a (3,0,2) model with 3 autocorrelation terms and
2 moving average terms.

The next step will be to see if I can improve on that with some different models
from the ACF and PACF graphs. Both the ACF and PACF look like they are somewhat
sinusoidal (PACF and ACF) and/or exponential (ACF) so an ARMA model is likely the
best choice.

I think I'll try a (2,1,2) model next and then a (3,1,2) model and see how they do
with AIC.

```{r}

AQres.212 <- arima(AQresiduals.ts, order=c(2,1,2), include.mean=FALSE)
summary(AQres.212)

AQres.312 <- arima(AQresiduals.ts, order=c(3,1,2), include.mean=FALSE)
summary(AQres.312)

AQres.303 <- arima(AQresiduals.ts, order=c(3,0,3), include.mean=FALSE)
summary(AQres.303)

AQres.313 <- arima(AQresiduals.ts, order=c(3,1,3), include.mean=FALSE)
summary(AQres.313)

AQres.304 <- arima(AQresiduals.ts, order=c(3,0,4), include.mean=FALSE)
summary(AQres.304)

```

After all those tests, the (3,0,3) model did best based on AIC:

(3,0,2) AIC: 3741.84
(3,1,3) AIC: 3738.84
(3,0,3) AIC: 3732.94

Next step is to look at the diagnostic plots.

```{r}

model1 = ggplot() + geom_point(aes(x=fitted(AQres.313), y=AQres.313$residuals)) + ggtitle("ARIMA (3,1,3)")
model2 = ggplot() + geom_point(aes(x=fitted(AQres.auto), y=AQres.auto$residuals)) + ggtitle("Auto (3,0,2)")
model3 = ggplot() + geom_point(aes(x=fitted(AQres.303), y=AQres.303$residuals)) + ggtitle("ARIMA (3,0,3)")
model4 = ggplot() + geom_point(aes(x=fitted(AQres.312), y=AQres.312$residuals)) + ggtitle("ARIMA (3,1,2)")

ggarrange(model1, model2, model3, model4, ncol=2, nrow=2)

```

No issues with the residuals vs. fitted.

```{r}

model1 = qplot(sample=AQres.313$residuals) + stat_qq_line(color="red") + ggtitle("ARIMA (3,1,3)")
model2 = qplot(sample=AQres.auto$residuals) + stat_qq_line(color="red") + ggtitle("Auto (3,0,2)")
model3 = qplot(sample=AQres.303$residuals) + stat_qq_line(color="red") + ggtitle("ARIMA (3,0,3)")
model4 = qplot(sample=AQres.312$residuals) + stat_qq_line(color="red") + ggtitle("ARIMA (3,1,2)")

ggarrange(model1, model2, model3, model4, ncol=2, nrow=2)

```

No problems with the Q-Q plots although the (3,0,3) looks the best as it doesn't
tail off as much at the upper end.

```{r}

# Diagnostics for independence
ggtsdiag(AQres.313,gof.lag=20)
ggtsdiag(AQres.auto,gof.lag=20)
ggtsdiag(AQres.303,gof.lag=20)
ggtsdiag(AQres.312,gof.lag=20)

```

The first three models look about the same with similar lags~=20. The last only
has about 11 lags of independence.

We'll continue with the best model so far: the ARIMA (3,0,3) model.

Look at the ACF and PACF of the residuals.

```{r}

AQres.303.acf <- ggAcf(AQres.303$residuals)
AQres.303.pacf <- ggPacf(AQres.303$residuals)
ggarrange(AQres.303.acf,AQres.303.pacf,nrow=2,ncol=1)

```

The spikes in the ACF and PACF at lag 11 are strange but the rest of the plots are
within the +/-0.1 bounds.

Next we need to look at the plot of the predicted vs. actual residuals.

```{r}

AQdata.fitted <- NO2.ts.lm6$fitted.values + fitted(AQres.auto)

ggplot() + geom_line(aes(x=time.NO2,y=NO2.ts[1:length(time.NO2)],color="True")) +
  geom_line(aes(x=time.NO2,y=AQdata.fitted,color="Fitted")) + xlab("Time") + 
  ylab("Daily NO2 level")

```

The Fitted plot looks pretty good, not as extreme at some points but generally 
trending close to the True value.

The next step is to use the model to predict the last 7 days that we didn't use
for training the model.

```{r}

AQdata.best.forecast <- forecast(AQres.303, h=7)

plot(AQdata.best.forecast)

next.7day.time <- c((length(dailyAQ.ts)-6):(length(dailyAQ.ts)))

dailyAQ.all.ts <- ts(dailyAQ)

# The test data frame
next.7day <- data.frame(time.NO2 = next.7day.time, NO2.GT. = NO2.ts[next.7day.time])

# The actual time series for the test period
next.7day.ts <- dailyAQ.ts[next.7day.time]
next.7day.ts <- ts(next.7day$NO2.GT.)

# Prediction for the next 7 days by ARIMA(3,0,3) model:
E_Y.pred <- predict(NO2.ts.lm6, newdata=next.7day)
e_t.pred <- forecast(AQres.303, h=7)
next.7day.prediction <- E_Y.pred + e_t.pred$mean

# MSE:
mean((next.7day.prediction-next.7day$NO2.GT.)^2)

time.predictions <- dailyAQ$Group.1[(length(time.NO2)+1):(length(time.NO2)+7)]
model1.predictions <- ggplot() + 
  geom_line(aes(x=time.predictions,y=next.7day$NO2.GT.),color="black") + 
  geom_line(aes(x=time.predictions,y=next.7day.prediction),color="red") + 
  geom_line(aes(x=time.predictions,y=E_Y.pred + e_t.pred$lower[,2]),
            color="red",linetype="dashed") + 
  geom_line(aes(x=time.predictions,y=E_Y.pred + e_t.pred$upper[,2]),
            color="red",linetype="dashed") +
  xlab("") + ylab("NO2") + 
  ggtitle("NO2 Trend + Seasonal Model + ARIMA of Residuals")

model1.predictions

```

That looks pretty good. The entire True line falls inside the lower and upper bounds
of the prediction model.

End of part 1

Part 2

Simulate a year's worth of NO2 data.

```{r}

# Simulate 365 days of NO2 with the best model
# auto.arima builds ARMA(2,1) model
set.seed(1)
auto.sim <- arima.sim(n=365, list(ar=c(AQres.303$coef[1],AQres.303$coef[2]),
                                      ma=c(AQres.303$coef[3])),
                        sd=sqrt(AQres.303$sigma2))

# Add mean predictions and plot simulation of NO2 level
next.365day.time <- c(1:365)
next.365day <- data.frame(time.NO2 = next.365day.time)

next.365day.predictions <- predict(NO2.ts.lm6, newdata=next.365day)

# plot simulated NO2 level
autoplot(ts(next.365day.predictions + auto.sim),xlab="Time",ylab="Simulated NO2 Level")

```

There are a couple of weird points where the simulated value falls below zero but
other than that it looks similar to the original, actual data set. I'm not sure if
this is for the next 365 days or for just a random 365-day interval.
